---
title: "Lab journal"
author: "by: Nina Sneijers"
bibliography: references.bib
editor_options: 
  markdown: 
    wrap: 72
---

```{r}
#loading packages
fpackage.check <- function(packages) {
    lapply(packages, FUN = function(x) {
        if (!require(x, character.only = TRUE)) {
            install.packages(x, dependencies = TRUE)
            library(x, character.only = TRUE)
        }
    })
}

fsave <- function(x, file = NULL, location = "./data/processed/") {
    ifelse(!dir.exists("data"), dir.create("data"), FALSE)
    ifelse(!dir.exists("data/processed"), dir.create("data/processed"), FALSE)
    if (is.null(file))
        file = deparse(substitute(x))
    datename <- substr(gsub("[:-]", "", Sys.time()), 1, 8)
    totalname <- paste(location, datename, file, ".rda", sep = "")
    save(x, file = totalname)  #need to fix if file is reloaded as input name, not as x. 
}

fload <- function(filename) {
    load(filename)
    get(ls()[ls() != "filename"])
}

fshowdf <- function(x, ...) {
    knitr::kable(x, digits = 2, "html", ...) %>%
        kableExtra::kable_styling(bootstrap_options = c("striped", "hover")) %>%
        kableExtra::scroll_box(width = "100%", height = "300px")
}

colorize <- function(x, color) {
    sprintf("<span style='color: %s;'>%s</span>", color, x)
}
packages = c("RSiena", "devtools", "igraph")
fpackage.check(packages)
#devtools::install_github('JochemTolsma/RsienaTwoStep', build_vignettes=TRUE)
packages = c("RsienaTwoStep")
fpackage.check(packages)
```
# Dataframe
Dataframe with the students characteristics for now: 
```{r}
load("C:/Users/ninas/OneDrive - Radboud Universiteit/RSCS/Social Networks/labjournalSN/data/All-ListedData_MRCP_Wave12_Ind.RData")
load("C:/Users/ninas/OneDrive - Radboud Universiteit/RSCS/Social Networks/labjournalSN/data/All-ListedData_MRCP_Wave12_Net.RData")

df1 <- as.data.frame(matrix(ncol=0, nrow=121))
#add student ID
df1$studentid <- studentids[[1]]
#adding gender
require(tidyverse)
df1$gender <- coalesce(gender[[1]][,1], gender[[1]][,2])
#add ethnicity 
df1$ethnicity <- coalesce(ethnic[[1]][,1], ethnic[[1]][,2])
#parentID 
view(parentedu)
df1$parent1edu <- coalesce(parentedu[[1]][,1], parentedu[[1]][,3])
df1$parent2edu <- coalesce(parentedu[[1]][,2], parentedu[[1]][,4])

#now, if I want to analyse parental education on a binary level I need to group them, so I will do that: 
library("car")
df1$parent1high <- car::recode(df1$parent1edu, "0:5 = 0; 6:10 = 1") #CHECK for 10 --> might be NA 
df1$parent2high <- car::recode(df1$parent2edu, "0:5 = 0; 6:10 = 1") 

df1$parentshigh <- with(df1, ifelse(parent1high == parent2high, parent1high, 0))
save(df1, file = "parentedudata.Rdata")
```

df <- df |>
    mutate(
    parent_norm = ifelse(conditie, a, ifelse(conditie2, b, NaN),
    parent_norm = ifelse (wat))
)

# Hanging NAs 
Dealing with NAs in my networkdata called 'hanging'. I am going to use the joiners/leavers method, so in the first wave it will be assumed that participants have send out no ties if they were absent, and if they joined later, the numbers from the first wave will be copied. 
```{r}
#so for wave 1: 
hangingw1 <- na_if(hanging[[1]][[1]], 9)
hangingw2 <- na_if(hanging[[1]][[2]], 9)

hangingw1_nona <- na_if(hanging[[1]][[1]], 9)
#checking if one misses in rows
rowhangingw1 <- rowSums(is.na(hangingw1_nona)) == nrow(hangingw1_nona)
rowhangingw1
#checking if one misses in cols
colhangingw1 <- colSums(is.na(hangingw1_nona)) == ncol(hangingw1_nona)
hangingw1_nona[which(rowhangingw1) , ] <- 0
hangingw1_nona

#CHANGE: for w2, I want the leavers to remain at the same value as they had in the previous wave, but I do not understand how to do this. Ask in class... for now, I will give them the 0 value. 
hangingw2
hangingw2_nona <- na_if(hanging[[1]][[2]], 9)
rowhangingw2 <- rowSums(is.na(hangingw2_nona)) == nrow(hangingw2_nona)
rowhangingw2
#checking if one misses in cols
colhangingw2 <- colSums(is.na(hangingw2_nona)) == ncol(hangingw2_nona)
hangingw2_nona[which(rowhangingw2) , ] <- 0
hangingw2_nona
```

# Factor analysis

```{r}
#factor analyse voor wave 1
head(T1socsat)
df_socsat1 <- as.data.frame(T1socsat[[1]])
df_socsat1$T1SOCSAT1_s <- scale(df_socsat1$T1SOCSAT1)
df_socsat1$T1SOCSAT2_s <- scale(df_socsat1$T1SOCSAT2)
df_socsat1$T1SOCSAT3_s <- scale(df_socsat1$T1SOCSAT3)
df_socsat1$T1SOCSAT4_s <- scale(df_socsat1$T1SOCSAT4)
df_socsat1$T1SOCSAT5_s <- scale(df_socsat1$T1SOCSAT5)

# Scale the data
socsat1_scaled <- scale(df_socsat1[,1:5])
library(psych)
fa1 <- fa(r = socsat1_scaled,
         nfactors = 4,
         rotate = "varimax")
summary(fa1)
fa1$loadings

#factor analysis voor wave 2
head(T2socsat)
df_socsat2 <- as.data.frame(T2socsat[[1]])
df_socsat1$T2SOCSAT1_s <- scale(df_socsat2$T2SOCSAT1)
df_socsat1$T2SOCSAT2_s <- scale(df_socsat2$T2SOCSAT2)
df_socsat1$T2SOCSAT3_s <- scale(df_socsat2$T2SOCSAT3)
df_socsat1$T2SOCSAT4_s <- scale(df_socsat2$T2SOCSAT4)
df_socsat1$T2SOCSAT5_s <- scale(df_socsat2$T2SOCSAT5)

# Scale the data
socsat2_scaled <- scale(df_socsat2[,1:5])
library(psych)
fa2 <- fa(r = socsat2_scaled,
         nfactors = 4,
         rotate = "varimax")
summary(fa2)
fa2$loadings
```

