---
title: "week4"
author: "by: Nina Sneijers"
bibliography: references.bib
version: "21-9-2026"
---
This week I have worked on graphs and webscraping. My R was messing with me, so the graphs can now be found in week 5, because for some reason in week 4, the file won't knit and in week 5, it will. 

```{r}
#loading packages
fpackage.check <- function(packages) {
    lapply(packages, FUN = function(x) {
        if (!require(x, character.only = TRUE)) {
            install.packages(x, dependencies = TRUE)
            library(x, character.only = TRUE)
        }
    })
}

fsave <- function(x, file = NULL, location = "./data/processed/") {
    ifelse(!dir.exists("data"), dir.create("data"), FALSE)
    ifelse(!dir.exists("data/processed"), dir.create("data/processed"), FALSE)
    if (is.null(file))
        file = deparse(substitute(x))
    datename <- substr(gsub("[:-]", "", Sys.time()), 1, 8)
    totalname <- paste(location, datename, file, ".rda", sep = "")
    save(x, file = totalname)  #need to fix if file is reloaded as input name, not as x. 
}

fload <- function(filename) {
    load(filename)
    get(ls()[ls() != "filename"])
}

fshowdf <- function(x, ...) {
    knitr::kable(x, digits = 2, "html", ...) %>%
        kableExtra::kable_styling(bootstrap_options = c("striped", "hover")) %>%
        kableExtra::scroll_box(width = "100%", height = "300px")
}

colorize <- function(x, color) {
    sprintf("<span style='color: %s;'>%s</span>", color, x)
}
packages = c("RSiena", "devtools", "igraph")
fpackage.check(packages)
#devtools::install_github('JochemTolsma/RsienaTwoStep', build_vignettes=TRUE)
packages = c("RsienaTwoStep")
fpackage.check(packages)
```

# Experimenting
```{r}
require(igraph)
g <- make_graph("Zachary")
plot(g)

#not the same pictures, but centres are the same
#centre of the graph is decided by amount of ties 
```

Description: 
Network exists of 34 people. Relations are non-directed (or is this the default of igraph?)
```{r}
gmat <- as_adjacency_matrix(g, type = "both", sparse = FALSE)
gmat
isSymmetric(gmat) 
#symmetric is true so the network is undirected
```

## Descriptives
```{r}
#size
# number of nodes
vcount(g)
# number of edges
ecount(g)

#degree
igraph::degree(g)
# hist(table(degree(g)), xlab='indegree', main= 'Histogram of indegree')

#transitivity
# be aware that directed graphs are considered as undirected. but g is undirected.
igraph::transitivity(g, type = c("localundirected"), isolates = c("NaN", "zero")) 

#betweenness
igraph::betweenness(g, directed = FALSE)

#1 and 34 should be stressed a bit more as they have higher betweenness. You could give them a colour with you palette or make them bigger


```

```{r}
igraph::dyad_census(g)

igraph::triad.census(g)
# I will use sna because it shows the names of the triads as well.
sna::triad.census(gmat)
unloadNamespace("sna")  #I will detach this package again, otherwise it will interfere with all kind of functions from igraph, and my students will hate me for that.
transitivity(
  g,
  type = c("globalundirected"),
  vids = NULL,
  weights = NULL,
  isolates = c("NaN", "zero")
)
```

visualisatie - we maken de grootte van de netwerken afhankelijk van iets anders: 
```{r}
# changing V(ertices)
# if you want to change the edges ((pijltjes)) use E 
V(g)$size = betweenness(g, normalized = T, directed = FALSE) * 60 + 12 #after some trial and error
plot(g, mode = "undirected")

#set the nodes a bit further apart
set.seed(2345)
l <- layout_with_mds(g)  #https://igraph.org/r/doc/layout_with_mds.html
plot(g, layout = l)


l  #let us take a look at the coordinates
l[1, 1] <- 4
l[34, 1] <- -3.5
plot(g, layout = l)
#this graph might make you interpret it like there is polarisation, fight, that 1 and 34 the network apart
```

Questions: is this data manipulation or not? Is it illustrating the story more clearly or not? Is it ethical? 
!!!! It is important to be open about this !!!
In this case you would mention in the article that you placed 1 and 34 on the outsides of your picture 

### Intermezzo: Panicking about NAs
```{r}
testm <- matrix(NA, nrow = 4, ncol = 4)
testm[1,1] <- 0 
testm[1,2] <- 1
testm[1,4] <- 1 
testm[3,1] <- 0 
testm[3,2] <- 1
testm[3,4] <- 1
testm[4,1] <- 1
testm[4,2] <- 0 
testm[4,4] <- 1 

#checking if one misses in rows
missrow <- rowSums(is.na(testm)) == nrow(testm)
#checking if one misses in cols
misscol <- colSums(is.na(testm)) == ncol(testm)
#checking if missing in both
misstotal <- which(missrow & misscol)

testm[which(missrow) , ] <- 0 
testm
```

# Webscraping
```{r}
# start with clean workspace
rm(list = ls())

# install.packages('data.table')
library(data.table)  # mainly for faster data handling
library(tidyverse)  # I assume you already installed this one!
# install.packages('httr') # we don't need this for now require(httr)
#install.packages("xml2")
require(xml2)
#install.packages("rvest")
require(rvest)
#install.packages("devtools")
require(devtools)
# Note we're doing something different here. We're installing a *latest* version directly from
# GitHub This is because the released version of this packages contains some errors!
devtools::install_github("jkeirstead/scholar")

require(scholar)

# define workdirectory, note the double *backslashes* if you're on windows setwd('/yourpathhere)'

url<- "https://nos.nl/zoeken?q=nprz&page=1"
nos_nprz <- read_html(url)
nos_nprz

#lijst <- nos_nprz %>% html_elements("main") %>% html_elements("ul") %>% html_elements("li")
```


```{r}
#titel <- nos |>

#lijst[1] %>% html_children %>% html_attr(name="href")
#lijst[1] %>% html_element("time") %>% html_text

#url <- paste0("https://nos.nl", artikel1) 
#tekst1 <- read_html(url)
#html_text2(tekst1)

#via element you go deeper into a page. you can also look into this via children 
```


------------------------------------------------------------------------

# References
:::
::::
