---
title: "week4"
author: "by: Nina Sneijers"
bibliography: references.bib
version: "21-9-2026"
---
```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
fpackage.check <- function(packages) {
    lapply(packages, FUN = function(x) {
        if (!require(x, character.only = TRUE)) {
            install.packages(x, dependencies = TRUE)
            library(x, character.only = TRUE)
        }
    })
}

fsave <- function(x, file = NULL, location = "./data/processed/") {
    ifelse(!dir.exists("data"), dir.create("data"), FALSE)
    ifelse(!dir.exists("data/processed"), dir.create("data/processed"), FALSE)
    if (is.null(file))
        file = deparse(substitute(x))
    datename <- substr(gsub("[:-]", "", Sys.time()), 1, 8)
    totalname <- paste(location, datename, file, ".rda", sep = "")
    save(x, file = totalname)  #need to fix if file is reloaded as input name, not as x. 
}

fload <- function(filename) {
    load(filename)
    get(ls()[ls() != "filename"])
}

fshowdf <- function(x, ...) {
    knitr::kable(x, digits = 2, "html", ...) %>%
        kableExtra::kable_styling(bootstrap_options = c("striped", "hover")) %>%
        kableExtra::scroll_box(width = "100%", height = "300px")
}

colorize <- function(x, color) {
    sprintf("<span style='color: %s;'>%s</span>", color, x)
}
packages = c("RSiena", "devtools", "igraph")
fpackage.check(packages)
#devtools::install_github('JochemTolsma/RsienaTwoStep', build_vignettes=TRUE)
packages = c("RsienaTwoStep")
fpackage.check(packages)
```

# Experimenting
```{r}
require(igraph)
g <- make_graph("Zachary")
plot(g)

#not the same pictures, but centres are the same
#centre of the graph is decided by amount of ties 
```

Description: 
Network exists of 34 people. Relations are non-directed (or is this the default of igraph?)
```{r}
gmat <- as_adjacency_matrix(g, type = "both", sparse = FALSE)
gmat
isSymmetric(gmat) 
#symmetric is true so the network is undirected
```

## Descriptives
```{r}
#size
# number of nodes
vcount(g)
# number of edges
ecount(g)

#degree
igraph::degree(g)
# hist(table(degree(g)), xlab='indegree', main= 'Histogram of indegree')

#transitivity
# be aware that directed graphs are considered as undirected. but g is undirected.
igraph::transitivity(g, type = c("localundirected"), isolates = c("NaN", "zero")) 

#betweenness
igraph::betweenness(g, directed = FALSE)

#1 and 34 should be stressed a bit more as they have higher betweenness. You could give them a colour with you palette or make them bigger


```

```{r}
igraph::dyad_census(g)

igraph::triad.census(g)
# I will use sna because it shows the names of the triads as well.
sna::triad.census(gmat)
unloadNamespace("sna")  #I will detach this package again, otherwise it will interfere with all kind of functions from igraph, and my students will hate me for that.
transitivity(
  g,
  type = c("globalundirected"),
  vids = NULL,
  weights = NULL,
  isolates = c("NaN", "zero")
)
```

visualisatie - we maken de grootte van de netwerken afhankelijk van iets anders: 
```{r}
# changing V(ertices)
# if you want to change the edges ((pijltjes)) use E 
V(g)$size = betweenness(g, normalized = T, directed = FALSE) * 60 + 12 #after some trial and error
plot(g, mode = "undirected")

#set the nodes a bit further apart
set.seed(2345)
l <- layout_with_mds(g)  #https://igraph.org/r/doc/layout_with_mds.html
plot(g, layout = l)


l  #let us take a look at the coordinates
l[1, 1] <- 4
l[34, 1] <- -3.5
plot(g, layout = l)
#this graph might make you interpret it like there is polarisation, fight, that 1 and 34 the network apart
```

Questions: is this data manipulation or not? Is it illustrating the story more clearly or not? Is it ethical? 
!!!! It is important to be open about this !!!
In this case you would mention in the article that you placed 1 and 34 on the outsides of your picture 

# Experimenting on own data
```{r}

load("C:\\Users\\ninas\\OneDrive - Radboud Universiteit\\RSCS\\Social Networks\\labjournalSN\\klooien.Rdata")
load("C:\\Users\\ninas\\OneDrive - Radboud Universiteit\\RSCS\\Social Networks\\labjournalSN\\data\\All-ListedData_MRCP_Wave12_Net.RData")

hanging[[1]][[1]]
hanging[[1]][[2]]
hanging[[1]][[3]]
#confusion! I thought there were two waves, what is the third about? 
hangingw1 <- na_if(hanging[[1]][[1]], 9)
hangingw2 <- na_if(hanging[[1]][[2]], 9)

hangingw1_nona <- na_if(hanging[[1]][[1]], 9)
#checking if one misses in rows
rowhangingw1 <- rowSums(is.na(hangingw1_nona)) == nrow(hangingw1_nona)
#checking if one misses in cols
colhangingw1 <- colSums(is.na(hangingw1_nona)) == ncol(hangingw1_nona)
hangingw1_nona[which(rowhangingw1) , ] <- 0
hangingw1_nona
```

Okay let's try something! 

```{r}
require(igraph)
gw1 <- igraph::graph_from_adjacency_matrix(hangingw1_nona, mode = "directed", weighted = NULL, diag = TRUE, add.colnames = NA,
    add.rownames = NA)
plot(gw1)
#ugly!
simplify(gw1)
#bit better... but still ugly
```

New attempt for wave 1 (looked at Jochem's plot):
```{r}
igraph::betweenness(gw1, directed = TRUE)
V(gw1)$size = betweenness(gw1, normalized = T, directed = TRUE) * 100 + 10
plot(gw1,
     vertex.color = "orange", # nice color for the vertices
     vertex.size = 5,  # we'll vertices a bit smaller
     vertex.frame.color = "gray",  # we'll put a gray frame around vertices
     vertex.label.color = "black",  # not that ugly blue color for the labels (names)
     vertex.label.cex = 0.5,  # make the label a bit smaller too
     vertex.label.dist = 0.9,  # we'll pull the labels a bit away from the vertices
     edge.curved = 0.2, # curved edges is always a nice touch
     edge.arrow.size = 0.1) # make arrow size (direction of edge) smaller
```


Let's compare to wave 2 and colour the men/women 
```{r}
hangingw2_nona <- na_if(hanging[[1]][[1]], 9)
#checking if one misses in rows
rowhangingw2 <- rowSums(is.na(hangingw2_nona)) == nrow(hangingw2_nona)
#checking if one misses in cols
colhangingw2 <- colSums(is.na(hangingw2_nona)) == ncol(hangingw2_nona)
hangingw2_nona[which(rowhangingw2) , ] <- 0

igraph::betweenness(gw2, directed = TRUE)
V(gw2)$size = betweenness(gw2, normalized = T, directed = TRUE) * 100 + 10
gw2 <- igraph::graph_from_adjacency_matrix(hangingw1_nona, mode = "directed", weighted = NULL, diag = TRUE, add.colnames = NA,
    add.rownames = NA)

kleurtje <- ifelse(gender_test==2, "orange", "purple")
plot(gw2,
     vertex.color = kleurtje, # nice color for the vertices
     vertex.size = 5,  # we'll vertices a bit smaller
     vertex.frame.color = "gray",  # we'll put a gray frame around vertices
     vertex.label.color = "black",  # not that ugly blue color for the labels (names)
     vertex.label.cex = 0.5,  # make the label a bit smaller too
     vertex.label.dist = 0.9,  # we'll pull the labels a bit away from the vertices
     edge.curved = 0.2, # curved edges is always a nice touch
     edge.arrow.size = 0.1)

gw2
```

### Intermezzo: Panicking about NAs
```{r}
testm <- matrix(NA, nrow = 4, ncol = 4)
testm[1,1] <- 0 
testm[1,2] <- 1
testm[1,4] <- 1 
testm[3,1] <- 0 
testm[3,2] <- 1
testm[3,4] <- 1
testm[4,1] <- 1
testm[4,2] <- 0 
testm[4,4] <- 1 

#checking if one misses in rows
missrow <- rowSums(is.na(testm)) == nrow(testm)
#checking if one misses in cols
misscol <- colSums(is.na(testm)) == ncol(testm)
#checking if missing in both
misstotal <- which(missrow & misscol)

testm[which(missrow) , ] <- 0 
testm
```

# Webscraping
```{r}
# start with clean workspace
rm(list = ls())

# install.packages('data.table')
library(data.table)  # mainly for faster data handling
library(tidyverse)  # I assume you already installed this one!
# install.packages('httr') # we don't need this for now require(httr)
#install.packages("xml2")
require(xml2)
#install.packages("rvest")
require(rvest)
#install.packages("devtools")
require(devtools)
# Note we're doing something different here. We're installing a *latest* version directly from
# GitHub This is because the released version of this packages contains some errors!
devtools::install_github("jkeirstead/scholar")

require(scholar)

# define workdirectory, note the double *backslashes* if you're on windows setwd('/yourpathhere)'

url<- "https://nos.nl/zoeken?q=nprz&page=1"
nos_nprz <- read_html(url)
nos_nprz

lijst <- nos_nprz %>% html_elements("main") %>% html_elements("ul") %>% html_elements("li")
```


```{r}
titel <- nos |>

lijst[1] %>% html_children %>% html_attr(name="href")
lijst[1] %>% html_element("time") %>% html_text

url <- paste0("https://nos.nl", artikel1) 
tekst1 <- read_html(url)
html_text2(tekst1)

#via element you go deeper into a page. you can also look into this via children 
```


------------------------------------------------------------------------

# References
:::
::::
